Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-04-23 lqg-strings.tex
just changed a minor thing
pushed the quantisation work to the main file

Rewriting eq(\ref{eqn:corrected-NG-wave-equation}) in terms of $\sigma$ derivatives and parameterizing $\sigma$ as (let $F^+ = 1 + g \Delta / 2 (- h)$)
\begin{equation} \label{eqn:sigma-parameterization}
	\sigma (q) = \frac{1}{T} \int _{0} ^{q} \diff E \frac{1}{(F^+)^2} = \frac{E (q)}{T (F^+)^2} ,
\end{equation}
we find the wave equation
\begin{equation} \label{eqn:simplified-modified-wave-eqn}
	(F^+)^2 \frac{\partial ^2 \Vec{X}}{\partial t^2} - \frac{\partial ^2 \Vec{X}}{\partial \sigma ^2} = 0 ,
\end{equation}
from which we can see that the correction factor modifies the speed of propagation of the wave in the string as
\begin{equation} \label{eqn:wave-speed-corrected}
	v = \frac{c}{F^+}.
\end{equation}

@@

In the same way we considered the quantum geometry inverse metric correction to the Nambu-Goto action, it is also natural to consider such correction to the Polyakov string. We start with the action(\ref{eqn:polyakov-gravity}) and proceed with the inverse area correction(\ref{eqn:inverse-metric-correction}), which yields our corrected Polyakov action
\begin{multline} \label{eqn:inverse-metric-Polyakov}
	S_{I A P} = - \frac{T}{2} \int \diff ^2 x \left( \sqrt{- g} - \frac{k \Delta}{2 \sqrt{- g}} \right) \times \\ \times g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} .
\end{multline}
The reason for considering this form and not one in which the inverse metric $g^{a b}$ also gets a correction term is that this form of a Polyakov inverse area-corrected action reduces nicely to the Nambu-Goto corrected action(\ref{eqn:modified-NG-action}) when the auxiliary metric $g_{a b}$ coincides with the induced metric $h_{a b} = \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu}$. The equations of motion for the auxiliary metric are completely unchanged by this correction, so they are
\begin{equation} \label{eqn:vary-wrt-g}
	\frac{\delta S_{I A P}}{\delta g^{c d}} = 0
\end{equation}
from which we obtain the stress-energy tensor for the worldsheet:
\begin{multline} \label{eqn:inverse-area-em-tensor}
	T_{c d} := \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} - \\ - \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} = 0 ,
\end{multline}
from which we can extract the metric $g_{c d}$ as
\begin{equation} \label{eqn:ws-metric}
	g_{c d} = 2 f \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} ,
\end{equation}
where the function $f$ is given by
\begin{equation} \label{eqn:inverse-area-f}
	\frac{1}{f} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} .
\end{equation}

@@

\begin{equation} \label{eqn:F-minus-exp}
	F^{-} = A \e ^{i (E \tau + p \sigma + \varphi)} , \ A > 0
\end{equation}
\begin{equation} \label{eqn:phi}
	\phi (\tau , \sigma) = \sqrt{\frac{k \Delta}{2 \left( 1 - A \e ^{i (E \tau + p \sigma + \varphi)} \right)}} .
\end{equation}
At first glance it seems problematic that we have a complex conformal factor $\phi (x)$, but the complex structure comes from our ansatz for the embedding fields, and when we choose either the real or imaginary part of the latter it will also pick the respective component of $\phi$. With this choice, we thus turned eq(\ref{eqn:X-eom-simplified}) into a Klein-Gordon equation
\begin{equation} \label{eqn:X-Klein-Gordon-eqn}
	\left( \partial _{\sigma} ^2 - \partial _{\tau} ^2 \right) X^{\mu} - \mu ^2 X^{\mu} = 0 .
\end{equation}

@@

In the following, we proceed to quantize the Klein-Gordon string via canonical covariant quantization, in a manner much similar to those found in \cite{Tong2009Lectures} and \cite{Green2012Superstring}. We turn the Poisson bracket relations into commutation relations between the now operator variables, $\{ , \} \mapsto i [ , ]$, yielding
\begin{align}
	\left[ \Hat{X}^{\mu} (\tau , \sigma) , \Hat{\Pi}^{\nu} (\tau , \sigma ') \right] & = i \delta (\sigma - \sigma ') \eta ^{\mu \nu} \label{eqn:commutator-X-Pi}                                      \\
	\left[ \Hat{X}^{\mu} , \Hat{X}^{\nu} \right]                                     & = \left[ \Hat{\Pi}^{\mu} , \Hat{\Pi}^{\nu} \right] = 0 \label{eqn:commutator-X-X-Pi-Pi}                         \\
	\left[ \Hat{a}^{\mu} _{n} , \Hat{a}^{\nu} _{m} \right]                           & = \left[ (\Hat{a}^{\mu} _{n})^{\dagger} , (\Hat{a}^{\nu} _{m})^{\dagger} \right] = 0 \label{eqn:commutator-a-a} \\
	\left[ \Hat{a}^{\mu} _{n} , (\Hat{a}^{\nu} _{m})^{\dagger} \right]               & = \eta ^{\mu \nu} \delta _{n , m} \label{eqn:commutator-a-a*}                                                   \\
	\left[ \Hat{x}^{\mu} , \Hat{x}^{\nu} \right]                                     & = \left[ \Hat{p}^{\mu} , \Hat{p}^{\nu} \right] = 0 \label{eqn:commutator-x-x-p-p}                               \\
	\left[ \Hat{x}^{\mu} , \Hat{p}^{\nu} \right]                                     & = i \eta ^{\mu \nu} . \label{eqn:commutator-x-p}
\end{align}

@@

As usual, by the appearance of the Minkowski metric in the non-zero commutator
\begin{equation} \label{eqn:ghosts}
	\Big[ \Hat{a}^{\mu} _{p} , (\Hat{a}^{\nu} _{k})^{\dagger} \Big] = \eta ^{\mu \nu} \delta _{p k} ,
\end{equation}
we have ghosts in the theory which must be dealt with. When translating the constraints as quantum operators, we require that they have vanishing matrix elements when sandwiched by physical states $\psi$ and $\phi$:
\begin{equation} \label{eqn:L-constraint-quantum-1}
	\bra{\phi} \Hat{L}_{p , k} \ket{\psi} = 0 \, \,
\end{equation}
\begin{equation} \label{eqn:L-constraint-quantum-2}
	\bra{\phi} \Hat{\widetilde{L}}_{p , k} \ket{\psi} = 0 .
\end{equation}

\section{Strings and Quantum Geometry}\label{sec:string-geometry}

>>>>

Rewriting eq(\ref{eqn:corrected-NG-wave-equation}) in terms of $\sigma$ derivatives and parameterizing $\sigma$ as (let $F^+ = 1 + g \Delta / 2 (- h)$)
\begin{equation} \label{eqn:sigma-parameterization}
	\sigma (q) = \frac{1}{T} \int _{0} ^{q} \diff E \frac{1}{(F^+)^2} = \frac{E (q)}{T (F^+)^2} ,
\end{equation}
we find the wave equation
\begin{equation} \label{eqn:simplified-modified-wave-eqn}
	(F^+)^2 \frac{\partial ^2 \Vec{X}}{\partial t^2} - \frac{\partial ^2 \Vec{X}}{\partial \sigma ^2} = 0 ,
\end{equation}
from which we can see that the correction factor modifies the speed of propagation of the wave in the string as
\begin{equation} \label{eqn:wave-speed-corrected}
	v = \frac{c}{F^+}.
\end{equation}
This is in fact an expected result for a effective theory emerging from a discretised background geometry, see\cite{Bodendorfer-LQG}.

@@

In the same way we considered the quantum geometry inverse metric correction to the Nambu-Goto action, it is also natural to consider such correction to the Polyakov string. We start with the action(\ref{eqn:polyakov-gravity}) and proceed with the inverse area correction(\ref{eqn:inverse-metric-correction}), which yields our corrected Polyakov action
\begin{multline} \label{eqn:inverse-metric-Polyakov}
	S_{I A P} = - \frac{T}{2} \int \diff ^2 x \left( \sqrt{- g} - \frac{k \Delta}{2 \sqrt{- g}} \right) \times \\ \times g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} .
\end{multline}
The reason for considering this form and not one in which the inverse metric $g^{a b}$ also gets a correction term is that this form of a Polyakov inverse area-corrected action reduces nicely to the Nambu-Goto corrected action(\ref{eqn:modified-NG-action}) when the auxiliary metric $g_{a b}$ coincides with the induced metric $h_{a b} = \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu}$. Also as we shall see, this action yields the exact same equation of motion as a more rigorous bimetric analysis, see appendix(\ref{app:bimetric-string}). The equations of motion for the auxiliary metric are completely unchanged by this correction, so they are
\begin{equation} \label{eqn:vary-wrt-g}
	\frac{\delta S_{I A P}}{\delta g^{c d}} = 0
\end{equation}
from which we obtain the stress-energy tensor for the worldsheet:
\begin{multline} \label{eqn:inverse-area-em-tensor}
	T_{c d} := \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} - \\ - \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} = 0 ,
\end{multline}
from which we can extract the metric $g_{c d}$ as
\begin{equation} \label{eqn:ws-metric}
	g_{c d} = 2 f \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} ,
\end{equation}
where the function $f$ is given by
\begin{equation} \label{eqn:inverse-area-f}
	\frac{1}{f} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu} .
\end{equation}

@@

\begin{equation} \label{eqn:F-minus-exp}
	F^{-} = A \e ^{i (E \tau + p \sigma + \varphi)} , \ A > 0
\end{equation}
\begin{equation} \label{eqn:phi}
	\phi (\tau , \sigma) = \sqrt{\frac{k \Delta}{2 \left( 1 - A \e ^{i (E \tau + p \sigma + \varphi)} \right)}} .
\end{equation}
At first glance it seems problematic that we have a complex conformal factor $\phi (x)$, but the complex structure comes from our ansatz for the embedding fields, and when we choose either the real or imaginary part of the latter it will also pick the respective component of $\phi$. With this choice, we thus turned eq(\ref{eqn:X-eom-simplified}) into a Klein-Gordon equation
\begin{equation} \label{eqn:X-Klein-Gordon-eqn}
	\left( \partial _{\sigma} ^2 - \partial _{\tau} ^2 \right) X^{\mu} - \mu ^2 X^{\mu} = 0 .
\end{equation}
The modification to the dispersion relation via the mass term is also an expected effect of a discrete background geometry as mentioned in\cite{Bodendorfer-LQG}.

@@

In the following, we proceed to quantize the Klein-Gordon string via canonical covariant quantization, in a manner much similar to those found in \cite{Tong2009Lectures} and \cite{Green2012Superstring}. We turn the Poisson bracket relations into commutation relations between the now operator variables, $[ , ] = i \{ , \}$, yielding
\begin{align}
	\left[ \Hat{X}^{\mu} (\tau , \sigma) , \Hat{\Pi}^{\nu} (\tau , \sigma ') \right] & = i \delta (\sigma - \sigma ') \eta ^{\mu \nu} \label{eqn:commutator-X-Pi}                                      \\
	\left[ \Hat{X}^{\mu} , \Hat{X}^{\nu} \right]                                     & = \left[ \Hat{\Pi}^{\mu} , \Hat{\Pi}^{\nu} \right] = 0 \label{eqn:commutator-X-X-Pi-Pi}                         \\
	\left[ \Hat{a}^{\mu} _{n} , \Hat{a}^{\nu} _{m} \right]                           & = \left[ (\Hat{a}^{\mu} _{n})^{\dagger} , (\Hat{a}^{\nu} _{m})^{\dagger} \right] = 0 \label{eqn:commutator-a-a} \\
	\left[ \Hat{a}^{\mu} _{n} , (\Hat{a}^{\nu} _{m})^{\dagger} \right]               & = \eta ^{\mu \nu} \delta _{n , m} \label{eqn:commutator-a-a*}                                                   \\
	\left[ \Hat{x}^{\mu} , \Hat{x}^{\nu} \right]                                     & = \left[ \Hat{p}^{\mu} , \Hat{p}^{\nu} \right] = 0 \label{eqn:commutator-x-x-p-p}                               \\
	\left[ \Hat{x}^{\mu} , \Hat{p}^{\nu} \right]                                     & = i \eta ^{\mu \nu} . \label{eqn:commutator-x-p}
\end{align}

@@

As usual, by the appearance of the Minkowski metric in the non-zero commutator
\begin{equation} \label{eqn:ghosts}
	\Big[ \Hat{a}^{\mu} _{p} , (\Hat{a}^{\nu} _{k})^{\dagger} \Big] = \eta ^{\mu \nu} \delta _{p k} ,
\end{equation}
we have ghosts in the theory which must be dealt with(\cite{Tong2009Lectures},\cite{Green2012Superstring}). When translating the constraints as quantum operators, we require that they have vanishing matrix elements when sandwiched by physical states $\psi$ and $\phi$:
\begin{equation} \label{eqn:L-constraint-quantum-1}
	\bra{\phi} \Hat{L}_{p , k} \ket{\psi} = 0 \, \,
\end{equation}
\begin{equation} \label{eqn:L-constraint-quantum-2}
	\bra{\phi} \Hat{\widetilde{L}}_{p , k} \ket{\psi} = 0 .
\end{equation}
\begin{equation} \label{eqn:L-constranit-quantum-sum}
	\bra{\phi} \sum _{p} \left( \left( \frac{p}{E_{p}} \right) ^2 - 1 \right) \Hat{\widetilde{L}}_{p , - p} \ket{\psi} = 0 .
\end{equation}

When translating the constraints as quantum operators, for $\Hat{L}_{p , k}$ we have no ambiguities since it's composed only of annihilation operators. As for $\Hat{\widetilde{L}}_{p , k}$, there is an ambiguity when $p = - k$, which manifests in the imposition of this constraint as
\begin{equation}
	\bra{\phi} \sum _{p} \left( \left( \left( \frac{p}{E_{p}} \right) ^2 - 1 \right) \Hat{\widetilde{L}}_{p , - p} \right) - a \ket{\psi} = 0 ,
\end{equation}
for some constant $a$\cite{Green2012Superstring}. This constant appears in the mass spectrum of the quantum string as in
\begin{multline}
	\Hat{M}^2 = \frac{1}{(4 \pi (\alpha ')^2)^2} \Bigg( (2 \pi \alpha ' \mu)^{2} \Hat{x}^{\mu} \Hat{x}_{\mu} - \\
	- \left( \sum _{n \neq 0} \left( \left( \frac{n}{E_{n}} \right) ^2 - 1 \right) (\Hat{a}^{\mu} _{n})^{\dagger} \Hat{a}^{\nu} _{n} \eta _{\mu \nu} - a \right) \Bigg) .
\end{multline}
The commutation algebra of constraints is inherited by the constraint algebra(\ref{eqn:L-bracket-0},\ref{eqn:L-bracket-1},\ref{eqn:L-bracket-2}), namely
\begin{equation} \label{eqn:L-commutator-0}
	\left[ \Hat{L}_{n , m} , \Hat{L}_{k , p} \right] = 0
\end{equation}
\begin{equation} \label{eqn:L-commutator-1}
	\left[ \Hat{L}_{n , m} , \Hat{\widetilde{L}}_{k , p} \right] = \Hat{L}_{m , k} \delta _{n , - p} + \Hat{L}_{n , k} \delta _{m , - p}
\end{equation}
\begin{equation} \label{eqn:L-commutator-2}
	\left[ \Hat{\widetilde{L}}_{n , m} , \Hat{\widetilde{L}}_{k , p} \right] = \Hat{\widetilde{L}}_{k , m} \delta _{n , - p} - \Hat{\widetilde{L}}_{n , p} \delta _{- m , k} .
\end{equation}
The only commutator with ordering ambiguities is(\ref{eqn:L-commutator-2}) when $m = - k$ and $p = -n$, thus we include anomalous terms\cite{Green2012Superstring}
\begin{multline} \label{eqn:L-commutator-anomalous}
	\left[ \Hat{\widetilde{L}}_{n , m} , \Hat{\widetilde{L}}_{k , p} \right] = \Hat{\widetilde{L}}_{k , m} \delta _{n , - p} - \Hat{\widetilde{L}}_{n , p} \delta _{- m , k} + \\
	+ C_{n} \delta _{n , -p} + D_{k} \delta _{- m , k} .
\end{multline}
By noting that $C_{- n} = - C_{n}$ and $D_{- n} = - D_{n}$ and using the Jacobi identity of the commutator, we conclude that $C_{n} = - D_{n} = C_{1} n$, where we can find the value of $C_{1}$ by calculating the vacuum expectation value of the commutator $[\Hat{\widetilde{L}}_{1 , k} , \Hat{\widetilde{L}}_{k , - 1}]$, which yields that $C_{1} = 0$, thus {\it the quantum algebra of our constraints has no anomalies}. This is a very important result, since it doesn't constrain the dimension of the background manifold to any particular value, so our model works just as well for $D = 4$ as for any other value {\it a priori}.

After analysing the mass-shell condition for the $n$-th excited state $\zeta _{\mu} (\Hat{a}^{\mu} _{n})^{\dagger} \ket{0 ; p}$\cite{Green2012Superstring}, under the assumption that the momentum wave-function is given by [justification needed]
\begin{equation} \label{eqn:p-wave-func}
	\psi (p) = \frac{\alpha '}{\mu} p^{2} ,
\end{equation}
we find that $\zeta _{\mu} p^{\mu} = 0$ and the momentum of this state is given by
\begin{equation} \label{eqn:nth-vector-mode-momentum}
	2 (4 \pi (\alpha ')^2)^2 p^2 = \frac{n^2}{n^2 + \mu ^2} - 1 - a .
\end{equation}
From this, for the absence of ghosts we then require that
\begin{equation} \label{eqn:a-condition-no-ghosts}
	a \geq \frac{d^2}{d^2 + \mu ^2} - 1 .
\end{equation}
In the boundary case, the $d$-th ($d = D - 1$) vector state is massless, while the other spin-1 states are massive and the ground state is a tachyon. Doing a similar analysis for the $d$-th spin-2 state $\zeta _{\mu \nu} (\Hat{a}^{\mu} _{d})^{\dagger} (\Hat{a}^{\nu} _{d})^{\dagger} \ket{0 ; p}$ with $\zeta _{\mu \nu} = \zeta _{\nu \mu}$ and $\zeta _{\ \mu} ^{\mu} = 0$ yields $\zeta _{\mu \nu} p^{\mu} = \zeta _{\mu \nu} p^{\nu} = 0$ and
\begin{equation} \label{eqn:a-conition-massless-spin-2}
	a = 2 \left( \frac{d^{2}}{d^{2} + \mu ^{2}} - 1 \right)
\end{equation}
for it to be massless. However, the quantity in parenthesis is negative, so doubling it makes it less than the minimum condition for the absense of ghosts(\ref{eqn:a-condition-no-ghosts}), thus we can't get massless spin-2 states. This is not really a problem because as pointed out in\cite{Bodendorfer-LQG}, a discrete background geometry modifies the dispersion relation at the effective level, so even tho at the fundamental level gravity would travel at $c$, in the effective limit it would appear \textquotedblleft massive", however with a very small mass. Thus, since we need $a$ to be at least as big as in(\ref{eqn:a-condition-no-ghosts}), we can choose $a > 0$ to also get rid of the ground state tachyon without the need to introduce supersymmetry or other unverified hypothesis.

The most natural value for $a$ is $a = 1$, which comes from the analysis of physical and spurious states\cite{Green2012Superstring} of the form
\begin{equation} \label{eqn:phys-spurious-state}
	\ket{\psi _{- 2 , n}} = \sum _{m} \Hat{\widetilde{L}}_{n , m} \ket{\widetilde{\chi}_{- 2 , m}} ,
\end{equation}
where $\ket{\widetilde{\chi}_{- 2 , m}}$ satisfies the constraints(\ref{eqn:L-constraint-quantum-1},\ref{eqn:L-constraint-quantum-2}) and also
\begin{multline} \label{eqn:spurious-basis-condition}
	\Bigg( \sum _{n} \left( \left( \left( \frac{n}{E_{n}} \right) ^2 - 1 \right) \Hat{\widetilde{L}}_{n , - n} \right) - \\
	- a + 2 - m \Bigg) \ket{\widetilde{\chi}_{- 2 , m}} = 0 .
\end{multline}
States of the form(\ref{eqn:phys-spurious-state}) are annihilated by $\Hat{\widetilde{L}}_{p , k}$ for $p \neq - k$, and for these cases we have
\begin{equation*}
	\sum _{n} \left( \left( \left( \frac{n}{E_{n}} \right) ^2 - 1 \right) \Hat{\widetilde{L}}_{- 1 , - n} \ket{\psi _{- 2 , n}} \right) =
\end{equation*}
\begin{multline*}
	= \sum _{n} \Bigg( \left( \left( \frac{n}{E_{n}} \right) ^2 - 1 \right) \Hat{\widetilde{L}}_{- 1 , - n} \times \\
	\times \sum _{m} \Hat{\widetilde{L}}_{n , m} \ket{\widetilde{\chi}_{- 2 , m}} \Bigg)  =
\end{multline*}
\begin{equation} \label{eqn:spurious-annihilation}
	= \sum _{n} \left( \left( \left( \frac{n}{E_{n}} \right) ^2 - 1 \right) \Hat{\widetilde{L}}_{n , - n} \right) \ket{\widetilde{\chi}_{- 2 , 1}} ,
\end{equation}
which vanishes for $a = 1$.

\section{Strings and Quantum Geometry} \label{sec:string-geometry}
